---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(pico)
```

# pico

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done
yet, or the repository is only intended to be a limited example, demo,
or
proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
<!-- badges: end -->

Create an R package with the minimum requirements.

## What

The aim of {pico} is to set up a skeleton directory with the _absolute minimum_ requirements for an R package.

This is a curio to demonstrate the bare bones of package requirements; it might be useful for intimidated R users who are creating packages for the very first time, perhaps to house frequently-used personal functions.

{pico} is absolutely not for 'serious' development of packages. There are many resources out there for proper package development that include all the additional stuff like documentation, testing, etc. For example:

* [Hilary Parker’s post](https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/) to write a package from scratch (beginner)
* [Tom Westlake’s update](https://r-mageddon.netlify.com/post/writing-an-r-package-from-scratch/) to Hilary’s post (beginner)
* [Emil Hvitfeldt’s post](https://www.hvitfeldt.me/blog/usethis-workflow-for-package-development/), focusing on {usethis} (beginner/intermediate)
* [Karl Broman’s site](https://kbroman.org/pkg_primer/), a primer for package development (intermediate)
* [Hadley Wickham’s book](http://r-pkgs.had.co.nz/) (intermediate/advanced)

## Install

Install {pico} from GitHub with help from {remotes}:

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("matt-dray/pico")
```

## Example

You can create a pico package with the `create()` function. To demonstrate, we can write the package 'picopkg' to a temporary folder.

```{r}
pico::create("picopkg", tempdir())
```

The minimum package content has now been created at the specified path. It comprises of:

* an `R/` directory with a `functions.R` file (which contains a dummy function called `say_hi()`)
* a `DESCRIPTION` file that contains only the name and version number of the package

```
picopkg/
├── R/
│   └── functions.R
└── DESCRIPTION
```

To add your functions, open the `picopkg/R/functions.R` file, add your function specifications and then save.

Now you can install your package (again with help from {remotes}):

```{r}
remotes::install_local(file.path(tempdir(), "picopkg"))
```

The functions from your pico package are now available to you. Here's the dummy function `say_hi()` in action:

```{r}
picopkg::say_hi("Matthew")
```
