---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

# Remove dummy package if already installed
if (any(as.data.frame(installed.packages())$Package == "mypkg")) {
 remove.packages("mypkg")
}

# Attach {pico}
library(pico)
```

# pico

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done
yet, or the repository is only intended to be a limited example, demo,
or
proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![R-CMD-check](https://github.com/matt-dray/pico/workflows/R-CMD-check/badge.svg)](https://github.com/matt-dray/pico/actions)
[![Codecov test coverage](https://codecov.io/gh/matt-dray/pico/branch/main/graph/badge.svg)](https://codecov.io/gh/matt-dray/pico?branch=main)
<!-- badges: end -->

Create a 'pico package': an R package with the minimum required structure and content.

## What

{pico} is a toy package that generates [the absolute bare-bones skeleton of an R package](http://kbroman.org/pkg_primer/pages/minimal.html). It may be a useful teaching aid to demystify the perceived complexity of R packages, or a quick-start to create a package of your often-used personal functions.

<details><summary>Click for (free) resources for 'proper' package-writing</summary>
* [Hilary Parker’s _Writing an R Package from Scratch_ post](https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/) (beginner)
* [Tom Westlake’s update to Hilary’s post](https://r-mageddon.netlify.com/post/writing-an-r-package-from-scratch/)  (beginner)
* [I wrote some slides about {usethis} for package development](https://www.rostrum.blog/2019/11/01/usethis/)
* [Emil Hvitfeldt’s {usethis} workflow](https://www.hvitfeldt.me/blog/usethis-workflow-for-package-development/) (beginner/intermediate)
* [Karl Broman’s _R Package Primer_ site](https://kbroman.org/pkg_primer/), a primer for package development (intermediate)
* [Hadley Wickham’s _R Packages_ book](http://r-pkgs.had.co.nz/) (intermediate/advanced)
</details>

## Install

Install {pico} from GitHub with help from {remotes}:

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("matt-dray/pico")
```

## Example

### Create

Use the `create()` function to generate a 'pico package' with the minimum required content. You supply a package `name` and a `dir`ectory where the package's folder will be written.

As a demonstration, here's how to create a pico package called {mypkg} in a temporary folder (you may want to put yours somewhere more convenient like `~/Documents` on macOS, for example):

```{r}
pico::create(
 name = "mypkg",
 dir = tempdir()
)
```

Note that the name will be checked against package-naming standards (i.e. it should contain alphanumeric characters or periods only and it cannot start with a number nor end with a period). The provided directory also will be checked for existence and---if it already contains a folder with the proposed name of your package---you'll be asked interactively if you want to overwrite it.

### Structure

The minimum package content has now been created at the specified path. The package structure looks like this:

```
mypkg/
├── R/
│   └── functions.R
└── DESCRIPTION
```

The `R/` directory contains a `functions.R` file, which is pre-filled with the dummy function `say_hi()` for demonstration purposes. The only other content is a special `DESCRIPTION` text-file that contains only the name and version number of the package.

### Modify

To add your own functions to the package:

1. Open the `mypkg/R/functions.R` file
2. Paste in your function specifications (i.e. like `add_one <- function(x) { x + 1 }`)
3. Save the file

You can create multiple `.R` files in `R/` to help you manage your functions.

### Install

How do you use your package? It lives on your computer (it's 'local'), so we can use the `install_local()` function from the {remotes} package (which we installed earlier). Pass the package's filepath to the function:

```{r}
remotes::install_local(
 path = file.path(tempdir(), "mypkg")
)
```

The package is now installed into your R package library and can be attached like any other package.

```{r}
library(mypkg)
```

So, for example, here's the dummy function `say_hi()` in action:

```{r}
say_hi("Matthew")
```

### Develop

You might want to add more functions later to `R/functions.R`. Paste them in and save the file. Now re-run `install_local()` with `force = TRUE` to overwrite the old version in the package library.

```{r}
remotes::install_local(
 path = file.path(tempdir(), "mypkg"),
 force = TRUE
)
```

Restart R and the new functions will be available from your package.

## Code of Conduct

Please note that the {pico} project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

```{r include=FALSE}
# Tidy up
unlink(file.path(tempdir(), "mypkg"), recursive = TRUE)
remove.packages("mypkg")
```
